Exploring the Soil Microbiome coupling with the Crete island Data Cube

### 2024-04-04 manuscript.tex
platforms and soil data

Soil ecosystems are the cornerstone of terrestrial habitats, biodiversity and henceforth human activities.
Soils are characterised by multiple properties; chemical, physical and biological that 
form complex interdependant interactions. Biodiversity of soils covers
all forms of life, fauna, flora, bacteria, archaea, fungi, viruses. 
Bacteria and archaea are considered major drivers for the functionality of soil.
They infuence and are infuenced by their environment and their community structure 
defies their macroscopic functionality.
Global soil microbiome studies have been employed to decipher soil microbiome compositions \cite{Delgado-Baquerizo2018, Labouyrie2023},
functions \cite{Bahram2018} and biogeography \cite{Martiny2006, Guerra2020}.
These results showed the remarkable diversity in soils yet there are blind spots \cite{Guerra2020}
and these sampling are sparce when considering samples per area density. One of most resolute
study is by \cite{Karimi2020} which exemplified the 
vast complexity of soil bacterial communuties and the requirement of dense samplings and isolated systems.

Islands are nature's labs \cite{Whittaker2017} because of their isolation and smaller scale.
Borrowing this paradigm, soil microbes \cite{Li2020} and mycorrhizal fungi \cite{Delavaux2021} studies
in islands show the benefits of using islands as models. Ultimately islands canr
set the ground to represent ecosystems with data and complex interactions \cite{Davies2016}.
This was also the case for the Island Sampling Day (ISD) project \cite{holm2024}
of the Genome Standards Consortium \cite{Field2011}
during the 18th workshop in June 2026 in Crete island, Greece. The goal was to "put standards into action"
in a soil microbiome survey with a dense sampling, 0.017 samples per km\textsuperscript{2}.
Hence, ISD is large scale study in the confined space of the island of Crete which 
is considered a miniature continent \cite{Vogiatzakis2008_crete}.

@@

Worldwide projects of microbiome studies have collected topsoil samples from Crete \cite{Vasar2022, Labouyrie2023, Bahram2018, Orgiazzi2018}.
Some have focused on soil fungi \cite{Mikryukov2023, Davison2021, Tedersoo2021} and other soil eukaryotes \cite{Aslani2022}.
These studies have

Koiliaris

@@

Spatial data

Integration of prior knowledge and 
Genome Standards Consortium \cite{Field2011}, FAIR data \cite{Wilkinson2016}, unrestricted \cite{Amann2019},
rich metadata and platforms of hosting open data like European Soil Data Centre \cite{Panagos2022}

Interactomes as a driver of community composition. Microbial interactions challenges \cite{Faust2021} cohesion?

In this study, we investigate

>>>>

Soil ecosystems are the cornerstone of terrestrial habitats, biodiversity and henceforth human activities.
Soils are characterised by multiple properties; chemical, physical and biological that 
form complex interdependant interactions. Biodiversity of soils covers
all forms of life, fauna, flora, bacteria, archaea, fungi, viruses. 
Bacteria and archaea are considered major drivers for the functionality of soil.
They infuence and are infuenced by their environment and their community structure 
defies their macroscopic functionality.
Global soil microbiome studies have been employed to decipher soil microbiome
compositions \cite{thompson2017a-communal, Delgado-Baquerizo2018, Labouyrie2023},
functions \cite{Bahram2018} and biogeography \cite{Martiny2006, Guerra2020}.
These results showed the remarkable diversity in soils yet there are blind spots \cite{Guerra2020}
and these sampling are sparce when considering samples per area density. One of most resolute
study is by \cite{Karimi2020} which exemplified the 
vast complexity of soil bacterial communuties and the requirement of dense samplings and isolated systems.

Islands are nature's labs \cite{Whittaker2017} because of their isolation and smaller scale.
Borrowing this paradigm, soil microbes \cite{Li2020} and mycorrhizal fungi \cite{Delavaux2021} studies
in islands show the benefits of using islands as models. Ultimately islands canr
set the ground to represent ecosystems with data and complex interactions \cite{Davies2016}.
This was also the case for the Island Sampling Day (ISD) project \cite{holm2024}
of the Genome Standards Consortium \cite{Field2011}
during the 18th workshop in June 2026 in Crete island, Greece. The goal was to "put standards into action"
in a soil microbiome survey with a dense sampling, 0.017 samples per km\textsuperscript{2}.
Hence, ISD is a large scale study in the confined space of the island of Crete which 
is considered a miniature continent \cite{Vogiatzakis2008_crete}.

@@

Worldwide projects of microbiome studies have collected one or two topsoil
samples from Crete \cite{Vasar2022, Labouyrie2023, Bahram2018, Orgiazzi2018}.
Some have focused on soil fungi \cite{Mikryukov2023, Davison2021, Tedersoo2021}
and other soil eukaryotes \cite{Aslani2022}.
The only thorough microbiome study of a soil ecosystem in Crete, to our knowledge,
is in the north west part of the island, the Koiliaris Critical Zone Observatory \cite{tsiknia2014}.
Apart from the soil microbiome, soil physical and chemical properties has been
investigaded by globan and european projects like the Forum of European Geological Surveys
(FOREGS) \cite{nerc19017}, the Geochemical Mapping of Agricultural and Grazing Land
Soil in Europe (GEMAS) \cite{REIMANN2018302} and the Soil Profile Analytical
Database for Europe (SPADE) \cite{Hiederer2006}.

Data and metadata of these samplings are stored in different databases, yet 
great effort from these distinct communities have led to establishing standards
to enable FAIR data \cite{Wilkinson2016}. For amplicon sequences the Genome Standards
Consortium \cite{Field2011} has established the MIMARKS \cite{yilmaz2011minimum}
standards and has been a advocate for open and unrestricted data \cite{Amann2019}.
For soil data, rich metadata and platforms of hosting open data exist, e.g the 
European Soil Data Centre \cite{Panagos2022} and the World Soil Information
Service (WoSIS) \cite{Batjes2024}. These platforms also provide maps.

Spatial and remote sensing data

@@

Integration of prior knowledge and

Interactomes have been characterised as a driver of community composition.
Microbial interactions challenges \cite{Faust2021} cohesion?

In this study, we integrade multiple types of data and methods to decipher hidden 
signals of the topsoil bacterial communities of Crete.
